---
title: "Analyse nLTT statistics"
author: "Richel Bilderbeek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analyse nLTTs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Abstract

This vignette shows how analyse the nLTT statistics.

I use this to check if `rJava` is installed correctly. 
If not, this will give an error:

```{r}
options(warn = 2)

library(wiritttes)
library(wiritttea)

file <- wiritttes::read_file(
  wiritttea::find_path("toy_example_1.RDa")
)
p <- wiritttes::get_posterior(file, sti = 1, ai = 1, pi = 1)

is_local_computer <- function()
{
  local_computer_names <- c(
    "fwn-biol-132-102", 
    "pc-157-104", 
    "lubuntu")
  return (sum(Sys.info()["nodename"] == local_computer_names) == 1)
}

is_local_computer()
```

## Setup

Loading this package

```{r cache=TRUE}
if (is_local_computer()) {
  # Read parameters
  parameters <- wiritttea::read_collected_parameters(
    "~/Peregrine/collect_files_parameters.csv"
  )
  
  # Read nLTT stats
  nltt_stats_raw <- wiritttea::read_collected_nltt_stats("~/Peregrine/collect_files_nltt_stats.csv")

  n_taxa <- wiritttea::read_collected_n_taxa("~/Peregrine/collect_files_n_taxa.csv")

  # No NA's in my dataset please
  nltt_stats <- nltt_stats_raw[!is.na(nltt_stats_raw$nltt_stat), ]
  
  # Merge nLTT statistics and parameters
  parameters$filename <- rownames(parameters)
  df <- merge(nltt_stats, parameters, by = "filename")
  df <- merge(n_taxa, df, by = "filename")
  df$scr <- as.factor(df$scr)
  df$mutation_rate <- as.factor(df$mutation_rate)
  df$sequence_length <- as.factor(df$sequence_length)
  df$sirg <- as.factor(df$sirg)
  df$siri <- as.factor(df$siri)
  df$erg <- as.factor(df$erg)
  df$eri <- as.factor(df$eri)
}
```

## Show the distribution of nLTT statistics

Here we have the bumps:

```{r fig.width = 7, fig.height = 7}
if (is_local_computer()) {
  
  ggplot2::ggplot(
  ) + ggplot2::facet_grid(siri ~ erg
  ) + ggplot2::geom_density(
    data = df,
    ggplot2::aes(
      x = nltt_stat,
      fill = scr
    )
    , alpha = 0.5
  )
}
```

Rampal wondered if this is caused by DNA sequence length. 
Here is short:

```{r fig.width = 7, fig.height = 7}
if (is_local_computer()) {
  
  ggplot2::ggplot(
  ) + ggplot2::facet_grid(siri ~ erg
  ) + ggplot2::geom_density(
    data = df[df$sequence_length == 1000, ],
    ggplot2::aes(
      x = nltt_stat,
      fill = scr
    )
    , alpha = 0.5
  )

}
```

Short has bumps.

Here is long:

```{r fig.width = 7, fig.height = 7}
if (is_local_computer()) {
  
  ggplot2::ggplot(
  ) + ggplot2::facet_grid(siri ~ erg
  ) + ggplot2::geom_density(
    data = df[df$sequence_length == 10000, ],
    ggplot2::aes(
      x = nltt_stat,
      fill = scr
    )
    , alpha = 0.5
  )

}
```

Long has bumps.

I predict it has to do with short trees. 

Here I show the number of taxa in the simulated trees,
for different speciation initiation rates, first as a histogram:

```{r fig.width = 7, fig.height = 7}
if (is_local_computer()) {

  ggplot2::ggplot(
      df, 
      ggplot2::aes(n_taxa, fill = siri)
  ) + ggplot2::geom_histogram()
}
```

here as a density plot:

```{r fig.width = 7, fig.height = 7}
if (is_local_computer()) {

  ggplot2::ggplot(
      df, 
      ggplot2::aes(n_taxa, fill = siri)
  ) + ggplot2::geom_density()
}
```

Or fancier, in a facet grid, as a histogram:

```{r fig.width = 7, fig.height = 7}
if (is_local_computer()) {

  ggplot2::ggplot(
  ) + ggplot2::facet_grid(siri ~ erg
  ) + ggplot2::geom_histogram(
    data = df,
    ggplot2::aes(
      n_taxa,
      fill = scr
    )
    , alpha = 0.5
  )
}
```

Or even fancier, in a facet grid, as a density plots:

```{r fig.width = 7, fig.height = 7}
if (is_local_computer()) {

  ggplot2::ggplot(
  ) + ggplot2::facet_grid(siri ~ erg
  ) + ggplot2::geom_density(
    data = df,
    ggplot2::aes(
      n_taxa,
      fill = scr
    )
    , alpha = 0.5
  )
}
```


```{r fig.width = 7, fig.height = 7}
if (is_local_computer()) {
  
  ggplot2::ggplot(
  ) + ggplot2::facet_grid(siri ~ erg
  ) + ggplot2::geom_density(
    data = df[df$sequence_length == 10000, ],
    ggplot2::aes(
      x = nltt_stat,
      fill = scr
    )
    , alpha = 0.5
  )

}
```

Zoom in on the bumps, focus on DNA sequence information:

```{r fig.width = 7, fig.height = 7}
if (is_local_computer()) {
  df_zoom <- df[df$sirg == 0.5 & df$eri == 0.0 ,]
  ggplot2::ggplot(
  ) + ggplot2::facet_grid(sequence_length ~ mutation_rate
  ) + ggplot2::geom_density(
    data = df_zoom,
    ggplot2::aes(
      x = nltt_stat,
      fill = scr
    )
    , alpha = 0.5
  )

}
```

Rescale this to a density of 100:

```{r fig.width = 7, fig.height = 7}
if (is_local_computer()) {
  ggplot2::ggplot(
  ) + ggplot2::facet_grid(sequence_length ~ mutation_rate
  ) + ggplot2::geom_density(
    data = df_zoom,
    ggplot2::aes(
      x = nltt_stat,
      fill = scr
    )
    , alpha = 0.5
  ) + ggplot2::scale_y_continuous(limits = c(0, 100))

}
```

Plot the number of taxa versus error:

```{r fig.width = 7, fig.height = 7}
if (is_local_computer()) {

  ggplot2::ggplot(
      df_zoom, 
      ggplot2::aes(x = n_taxa, y = nltt_stat)
  ) + ggplot2::geom_point(
  ) + ggplot2::geom_smooth(method = "lm")
  

}
```

